---
title: "Intro to Data Visualization"
author: "Clarissa Reyes"
date: "2023-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r}
library(dplyr)
library(readr)
library(tidyr)
library(forcats) #order factors better
library(ggplot2)
library(leaflet) #maps
library(DT) #data table
library(scales)
```

## Read data
```{r}
escapement <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e")
```

## Explore data
```{r}
colnames(escapement)
glimpse(escapement)
unique(escapement$Species)
head(escapement)
range(escapement$sampleDate)

```
## Prepare data
```{r}
annual_esc <- escapement %>% 
  separate(sampleDate, c("Year","Month", "Day"), sep = "-") %>%  #separate date elements
  mutate(Year = as.numeric(Year)) %>% #changing year from character to numeric
  group_by(Species, SASAP.Region, Year) %>% 
  summarise(escape = sum(DailyCount)) %>% #calculate total sum
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

unique(annual_esc$Species)
head(annual_esc)
```

## Plotting with `ggplot`

### Essential components
```{r}
#Option1
ggplot(data = annual_esc,
      #mapping aesthetics
       aes(x = Species, 
           y = escape))+
  geom_col() #bar plot
 
#Option2 
ggplot(data = annual_esc)+
  geom_col(aes(x = Species, y = escape))

#Option3
ggplot()+
  geom_col(data = annual_esc,
           aes(x = Species,
               y = escape))
```

### Diff geoms

#### Boxplot

```{r}
annual_esc %>% 
  filter(Year==1974, #don't need quotes bc its a number
         Species %in% c("Chum","Pink")) %>% 
  ggplot(aes(x = Species,
             y = escape))+
  geom_boxplot()

```
#### Line and point
```{r}
annual_esc %>% 
  filter(Species == "Sockeye", #don't need quotes bc its a number
         SASAP.Region == "Bristol Bay") %>% 
  ggplot(aes(x = Year,
             y = escape))+
  geom_line()+
  geom_point()

```
### Customizing plot

Blue inside aes()
```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escape,
           fill = "blue"))+
  geom_col()
```

Blue outside aes()
```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escape))+ #everything inside aes addresses one variable 
  geom_col(fill = "blue") 
```
Adding color by region
```{r}
ggplot(annual_esc,
       aes(x = Species,
           y = escape,
           fill = SASAP.Region))+  
  geom_col() 
```
### Faceted plot

```{r}
annual_esc_2000 <- annual_esc %>% 
  filter(Year %in% 2000:2016)

```

###ggplot themes
ggplot(an)
```{r}
ggplot(annual_esc_2000,
       aes( x= Year,
            y = escape,
            color = Species))+
  geom_line()+
  geom_point()+
  facet_wrap(~SASAP.Region,  #~ calls variable within our dataset
             scales = "free_y")+
  theme_light()+
  ylab("Escapement")+
  ggtitle("Annual Salmon Escapement by Region")+
  theme(legend.position = "bottom",
        legend.title = element_blank())
```
### Saved Themes
```{r}
my_theme <- theme_light()+
    theme(legend.position = "bottom",
        legend.title = element_blank())
```

### Exercise
```{r}
ggplot(annual_esc_2000,
       aes( x= Year,
            y = escape,
            color = Species))+
  geom_line()+
  geom_point()+
  facet_wrap(~SASAP.Region,  #~ calls variable within our dataset
             scales = "free_y")+
  ylab("Escapement")+
  my_theme+
  theme(axis.text.x = element_text(angle=45,
                                   vjust = 0.5)) #vertical position of x label

annual_region_plot <- ggplot(annual_esc_2000,
       aes( x= Year,
            y = escape,
            color = Species))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(2000,2016,2))+
  scale_y_continuous(labels = comma)+ #y scale has commas
  facet_wrap(~SASAP.Region,  #~ calls variable within our dataset
             scales = "free_y")+
  ylab("Escapement")+
  my_theme+
  theme(axis.text.x = element_text(angle=45,
                                   vjust = 0.5)) #vertical position of x label
```

### Saving plots
```{r}

ggsave("figs/annual_esc_region.png", annual_region_plot, width = 12, height = 8, units = "in") 
```

## Maps with `leaflet`
Interactive map that plots regions

### Subset of locations
```{r}
location <- escapement %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
glimpse(location)
```

### Create map 
Leaflet uses pipe, ggplot uses +
```{r}
leaflet(location) %>%
  addTiles() %>% # from open street map
  addMarkers(lng = ~Longitude,
             lat = ~Latitude,
             popup = ~Location)
```

### Map with diff basemap

```{r}
leaflet(locations) %>%
  # using a specific basemap
    addWMSTiles(
        "https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?request=getmap&service=wms&BBOX=-90,-180,90,360&crs=EPSG:4326&format=image/jpeg&layers=gebco_latest&width=1200&height=600&version=1.3.0",
        layers = 'GEBCO_LATEST',
        attribution = "Imagery reproduced from the GEBCO_2022 Grid, WMS 1.3.0 GetMap, www.gebco.net"
    ) %>%
  # marker details
    addCircleMarkers( 
        lng = ~ Longitude,
        lat = ~ Latitude,
        popup = ~ Location,
        radius = 5,
        # set fill properties
        fillColor = "salmon",
        fillOpacity = 1,
        # set stroke properties
        stroke = T,
        weight = 0.5,
        color = "white",
        opacity = 1
    )
```


