---
title: "Intro to sf"
author: "Clarissa Reyes"
date: "2023-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r, message=FALSE}
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap) #used to add base maps to static maps 
library(dplyr)
library(readr)

```

# Load and explore AK Regions data
```{r}
ak_regions <- read_sf("data/ak_regions_simp.shp")

plot(ak_regions)
class(ak_regions) #classes incl sf AND dataframe
glimpse(ak_regions)
```

# Transform CRS of `ak_regions`
```{r}
st_crs(ak_regions) #geocrs is unprojected, not meant to be plotted on a map

ak_regions_3338 <- ak_regions %>% 
  st_transform(crs = 3338) #tell which EPSG code you want to do

st_crs(ak_regions_3338) #now in projected crs
plot(ak_regions_3338)

```
# Notes
crs (coord ref system)
unprojected: geographic coordinate system (gcs)
- defines where something is positioned on the globe #D sphere, angular - typically degrees

projected: has lat/long, represented as 2D- typically in linear unit, ie meters

every map of the earth is distorted in some way
EPSG codes for different projections
prj file gives details about the projection you've started

4236 is commonly used because it has coordinates for the whole world
3857 used in google, arcGIS, etc

# sf and tidyverse
```{r}
colnames(ak_regions_3338)

ak_regions_3338 %>% 
  select(region) %>%  #choose what you want to keep; geometry will remain ("sticky") bc the data itself is .shp
  filter(region == "Yukon")
```

# spatial joins
```{r}
pop <- read_csv("data/alaska_population.csv")

#explore data
View(pop) #is in EPSG 3338 
class(pop) #no sf class - no geometry col


pop_4326 <- st_as_sf(pop, #changes to spatial projection
                     coords = c("lng","lat"),
                     crs = 4236,
                     remove = FALSE) #do not remov long and lat, but can bc because geometry col now exists
class(pop_4326) #now has sf class

pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within) #wanting to know pops within those regions
#can't join because crs of both datasets don't match

pop_3338 <- st_transform(pop_4326, crs= 3338) #changed from 4326 to 3338
st_crs(pop_3338) # can check projections

pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within) 
glimpse(pop_joined)
```

# Wrangling
```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>%  # want df in order to use tidyverse functions
  group_by(region) %>% 
  summarise(total_pop = sum(population))
head(pop_region)

class(pop_region) #still shows df, but need it to be spatial for mapping

pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")
```
# Plot
```{r}
plot(pop_region_3338["total_pop"]) #square brackets allow you to index: access diff values/variables within df

```

# New plot by mgmt
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop))

plot(pop_mgmt_3338["total_pop"]) # lost some of the AK regions

pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop), do_union = FALSE) #don't want to lose the geometries 
View(pop_mgmt_3338)

write_sf(pop_region_3338, "data/ak_regions_population.shp") #save
```
# Visualize with ggplot
```{r}
ggplot(pop_region_3338) +
    geom_sf(aes(fill = total_pop))+ 
    labs(fill = "Total Population")+ 
    scale_fill_viridis_c(labels = comma)+ #colorblind 
    theme_bw()
```

# Rivers spatial data
```{r}
rivers_3338 <- read_sf("data/ak_rivers_simp.shp")
st_crs(rivers_3338) #see crs - using same projection as ak_pop

ggplot()+
  #image layers
  geom_sf(pop_region_3338, mapping = aes(fill = log10(total_pop)))+ #one data obj here in geom_sf layer
  geom_sf(pop_3338, mapping = aes(), color = "red4", size = 0.5)+ #another data obj #keep aes blank bc using other mappting aes #color has nothing to do with data so should be outside of aes()
  geom_sf(rivers_3338, mapping = aes(linewidth = StrOrder))+ 
    scale_linewidth(range = c(0.05,0.5), #changes scale of line widths
                    guide = "none")+ #removes stream order legend 
  #customization
  scale_fill_viridis_c(labels = comma)+ #defines legend numbers to take away from scientific notation
  labs(fill = "Total Population",
       title = "Alaska populations")+ #fill corresponds to total_pop variable 
  theme(
    plot.title = element_text(hjust = 0.5)
    )
```   
# Visualize `sf` objs with `leaflet`
You have to provide projection AND unprojected

ggplot uses + (to layer, so next level overrides)
leaflet uses pipes ("and then")
leafletCRS allows us to create a custom crs


```{r}

epsg3338 <- leaflet::leafletCRS( #custom crs to create projected
    crsClass = "L.Proj.CRS",
    code = "EPSG:3338",
    proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs", #proj4 definition required from EPSG
    resolutions = 2 ^ (16:7)
)

pop_region_4326 <- pop_region_3338 %>%  #unprojected
  st_transform(crs=4326) 

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>% 
  addPolygons(data = pop_region_4326,
              fillColor = "green",
              weight = 1,
              fillOpacity = 2) #line thickness 
m
```   




