---
title: "RMarkdown Analysis"
author: "Clarissa Reyes"
date: "2023-04-03"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction
Data set contains North Pole Environmental Observatory Bottle Chemistry data.
Data downloaded on April 3, 2023 at [this link](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)


### Setup
readr package is part of tidyverse
```{r}
library(readr) #load packages 
```

### Read data
reader:: specifies that the function comes from reader package 
```{r read data, warning=FALSE, message=FALSE}
bg_chem <- readr::read_csv("data/BGchem2008data.csv") #faster for import, good with deal w factors

```

#### Data exploration
eval: run or don't run
echo: show or don't show code in knitted product
```{r, eval=FALSE, echo=FALSE}
#names of columns
colnames(bg_chem)

#data structure
str(bg_chem)

#summary
summary(bg_chem)

#view in a tab
View(bg_chem)

#unique values in a column
unique(bg_chem$Date)
```

## Analysis

### Calculate summary stats
```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

### Calc mean redfield ratio
Redfield ratio = 16:1

```{r}
ratio <- (nitrate + nitrite + amm)/phos
```

### Plot
```{r}
plot(bg_chem$P, bg_chem$NO3 + bg_chem$NO2 + bg_chem$NH4)
```

## Conclusion

Can call an object value in-text -- this is good for automatic updating of values, ie.

The redfield ratio is `r round(ratio)`